Coexistence with varying environmental filtering intensity

### 2018-06-11 env_intensity.Rmd
:sparkles: Split chunks + multiple trait gen
Split chunks to avoid cluttering console. Add function to generate
several traits with various correlation coefficients with one trait
following the environment. Change FD plots.

---
title: "Coexistence with varying environmental filtering intensity"
author: "XX"
date: "5 juin 2018"
output: pdf_document
---

@@

```{r}

devtools::load_all()
#### Keep stable for all
patches  <- 25   # Number of patches
species  <- 25   # Number of species
time     <- 150	 # Length of model run (generations)
initpop  <- 50   # initial population size
n_traits <- 10    # number of traits
k        <- 1.45 # Niche height
width    <- seq(10, 0.25, length.out = patches) # Niche breadth

@@

# Generate traits
set.seed(1)
traits <- matrix(runif(species*n_traits, min(env), max(env)),
                 species,
                 n_traits)
colnames(traits) <- paste0("trait", seq(n_traits))
rownames(traits) <- paste0("species", seq(species))

# Number of trait influencing growth or competition
trait_types = t(data.frame(prop1 = c(rep("R", 8), rep("A", 2)),
                           prop2 = c(rep("R", 5), rep("A", 5)),
                           prop3 = c(rep("R", 2), rep("A", 8))))
colnames(trait_types) = colnames(traits)

multiple_traits <- apply(trait_types, 1,
                         function(x){
                             multiple_traits_types = multigen(traits = traits, trait_type = x,
                                                              env = env, time = time, species = species,
                                                              patches = patches, composition = composition,
                                                              A = A, d = d, k = k, width = width)
                         })
names(multiple_traits) <- rownames(trait_types)

plot_patches_list <- lapply(names(multiple_traits),
                            function(x){
                                plot_patch(multiple_traits[[x]]$compo, 15, 150) +
                                    labs(subtitle = x)
                            })

@@

sp_rich_final <- as.data.frame(sp_rich_final)
colnames(sp_rich_final) <- names(multiple_traits)
sp_rich_final$width <- width

@@

# Compute FD metrics (richness, dispersion, eveness)
timestep = 10
FD_prop <- lapply(names(multiple_traits),
                  function(x){
                      patch_sp_timestep <- multiple_traits[[x]]$compo[,,timestep]
                      ab_rel = patch_sp_timestep / rowSums(patch_sp_timestep)
                      ab_rel = ab_rel[!is.na(rowSums(ab_rel)), ]
                      FD_patches <- sapply(seq_len(nrow(ab_rel)), function(y){
                          ab_patch = ab_rel[y, ][ab_rel[y, ] > 0] # Removal of species with null abundance
                          if(length(ab_patch) > 1){ # Patch must contain at least 2 species
                              FD_patch = dbFD(traits[rownames(traits) %in%
                                                         names(ab_patch), ],
                                              a = ab_patch,
                                              w.abun = TRUE)
                              data.frame(prop = x,
                                         patch = rownames(ab_rel[y, , drop = FALSE]),
                                         frich = FD_patch$FRic,
                                         fdis = FD_patch$FDis,
                                         feve = FD_patch$FEve)
                          }
                      })
                      FD_patches <- do.call(rbind, FD_patches)
                  })
FD_prop <- do.call(rbind, FD_prop)

@@

@@

# Null models: resample species within each patch (NM2)
# FD null for each proportion case
nb_perm = 99#99
timestep = 10
FD_null_prop <- lapply(names(multiple_traits),
                       function(x){
                           # Null model 2 (Bernard-Verdier et al. 2012) => shuffling abundances of species within patches
                           null_matrices <- replicate(nb_perm,
                                                      t(apply(multiple_traits[[x]]$compo[, , timestep], 1, sample)),
                                                      simplify = FALSE)
                           # Convert to relative abundances
                           null_matrices <- lapply(null_matrices, function(y) y/rowSums(y))
                           # Add species in column names
                           null_matrices = lapply(null_matrices, as.data.frame)
                           null_matrices = lapply(null_matrices, setNames, paste0("species", seq(species)))
                           # Remove empty rows in null matrices
                           null_matrices = lapply(null_matrices, function(x) x[!is.na(rowSums(x)), ])

@@

FD_prop <- left_join(FD_prop, FD_null_prop, by = c("prop", "patch"))
FD_prop$frich_SES <- (FD_prop$frich - FD_prop$frich_mean) / FD_prop$frich_sd
FD_prop$fdis_SES <- (FD_prop$fdis - FD_prop$fdis_mean) / FD_prop$fdis_sd
FD_prop$feve_SES <- (FD_prop$feve - FD_prop$feve_mean) / FD_prop$feve_sd

# Plot
plot_prop_frich_SES <- ggplot(FD_prop, aes(1/width, frich_SES, color = as.factor(prop))) +
    geom_line(size = 1) +
    scale_color_manual(values = viridis::viridis(3)) +
    xlab("Environmental intensity (1/width)") +
    ylab("Patch functional richness SES") +
    theme(legend.position = "none")

>>>>

---
title: "Coexistence with varying environmental filtering intensity"
author: "XX"
date: "5 juin 2018"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

@@

## Generate several traits with various scenarios

### Moderately correlated traits
```{r}
generate_cor_traits = function(number_patches, number_other = 9, cor_coef = 0.7) {

    # Initial Trait
    x1 = seq(1, number_patches)

    cov_mat = matrix(c(1, cor_coef, 0, sqrt(1 - cor_coef^2)), nrow = 2)

    n_other = number_other

    res = lapply(seq(n_other), function(x) {
        x2 = runif(number_patches, 1, number_patches)

        cov_mat %*% matrix(c(x1, x2), nrow = 2, byrow = TRUE)
    })

    res = do.call(rbind, lapply(res, function(x) x[2,]))
    res = rbind(x1, res)

    t(res)
}

devtools::load_all()
# Initial parameters -----------------------------------------------------------
patches  <- 25   # Number of patches
species  <- 25   # Number of species
time     <- 150	 # Length of model run (generations)
initpop  <- 50   # initial population size
n_traits <- 4    # number of traits
k        <- 1.45 # Niche height
width    <- seq(10, 10, length.out = patches) # Niche breadth

@@

# Generate traits --------------------------------------------------------------
set.seed(1)
traits <- generate_cor_traits(patches, n_traits - 1, cor_coef = 0.7)
colnames(traits) <- paste0("trait", seq(n_traits))
rownames(traits) <- paste0("species", seq(species))

# Number of trait influencing growth or competition
trait_types = t(data.frame(prop_8_2 = c(rep("R", round(0.8*n_traits)),
                                        rep("A", n_traits - round(0.8*n_traits))),
                           prop_5_5 = c(rep("R", round(0.5*n_traits)),
                                        rep("A", n_traits - round(0.5*n_traits))),
                           prop_2_8 = c(rep("R", round(0.2*n_traits)),
                                        rep("A", n_traits - round(0.2*n_traits)))))
colnames(trait_types) = colnames(traits)

# Simulations ------------------------------------------------------------------
multiple_traits <- apply(trait_types, 1,
                         function(x){
                             multiple_traits_types = multigen(traits = traits, trait_type = x,
                                                              env = env, time = time, species = species,
                                                              patches = patches, composition = composition,
                                                              A = A, d = d, k = k, width = width)
                         })
names(multiple_traits) <- rownames(trait_types)

# Plots ------------------------------------------------------------------------

# Plot individual patches 
plot_patches_list <- lapply(names(multiple_traits),
                            function(x){
                                plot_patch(multiple_traits[[x]]$compo, 15, 150) +
                                    labs(subtitle = x)
                            })

@@

sp_rich_final <- as.data.frame(sp_rich_final)
colnames(sp_rich_final) <- names(multiple_traits)
sp_rich_final <- tibble::rownames_to_column(sp_rich_final, "patch")
sp_rich_final$patch <- as.numeric(gsub("patches", "", sp_rich_final$patch))
sp_rich_final$width <- width

@@

plot_prop_sp_rich_patch <- ggplot(sp_rich_final, aes(patch, richness, color = prop_type)) +
    geom_line(size = 1) +
    scale_color_manual(values = viridis::viridis(3)) +
    xlab("Patch Index") +
    ylab("Patch richness")

```{r FD_compute}
# Compute FD metrics -----------------------------------------------------------
# Compute FD metrics (richness, dispersion, eveness)
timestep = 150
FD_prop <- lapply(names(multiple_traits),
                  function(x) {

                      patch_sp_timestep <- multiple_traits[[x]]$compo[,,timestep]
                      ab_rel = patch_sp_timestep / rowSums(patch_sp_timestep)
                      ab_rel = ab_rel[!is.na(rowSums(ab_rel)), ]

                      # Compute Functional Diversity at Patch Scale
                      FD_patches <- sapply(seq_len(nrow(ab_rel)), function(y){
                          ab_patch = ab_rel[y, ][ab_rel[y, ] > 0] # Removal of species with null abundance
                          if (length(ab_patch) > 1) { # Patch must contain at least 2 species
                              FD_patch = dbFD(traits[rownames(traits) %in%
                                                         names(ab_patch), ],
                                              a = ab_patch,
                                              w.abun = TRUE)

                              data.frame(prop = x,
                                         patch = rownames(ab_rel[y, , drop = FALSE]),
                                         frich = FD_patch$FRic,
                                         fdis = FD_patch$FDis,
                                         feve = FD_patch$FEve)
                          }
                      }, simplify = FALSE)
                      bind_rows(FD_patches)
                  })
FD_prop <- bind_rows(FD_prop)

@@

FD_prop$patch <- as.numeric(gsub("patches", "", FD_prop$patch))

FD_melt <- tidyr::gather(FD_prop, "fd_metric", "fd_value", frich:feve)

@@

plot_fd_patch <- ggplot(FD_melt, aes(patch, fd_value, color = prop)) +
    geom_line() +
    facet_grid(fd_metric~., scales = "free_y") +
    scale_color_manual("Proportion of traits\ncontributing to R A",
                       values = viridis::viridis(3)) +
    labs(x = "Patch Index",
         y = "FD value")
```

```{r null_models}
# Null models ------------------------------------------------------------------
# Null models: resample species within each patch (NM2)
# FD null for each proportion case
nb_perm = 99#99
timestep = 10
FD_null_prop <- lapply(names(multiple_traits),
                       function(x){
                           # Null model 2 (Bernard-Verdier et al. 2012) => shuffling abundances of species within patches
                           null_matrices <- replicate(nb_perm,
                                                      t(apply(multiple_traits[[x]]$compo[, , timestep], 1, sample)),
                                                      simplify = FALSE)
                           # Convert to relative abundances
                           null_matrices <- lapply(null_matrices, function(y) y/rowSums(y))
                           # Add species in column names
                           null_matrices = lapply(null_matrices, as.data.frame)
                           null_matrices = lapply(null_matrices, setNames, paste0("species", seq(species)))
                           # Remove empty rows in null matrices
                           null_matrices = lapply(null_matrices, function(x) x[!is.na(rowSums(x)), ])

@@

# Compute FD SES ---------------------------------------------------------------
FD_prop <- left_join(FD_prop, FD_null_prop, by = c("prop", "patch"))
FD_prop$frich_SES <- (FD_prop$frich - FD_prop$frich_mean) / FD_prop$frich_sd
FD_prop$fdis_SES <- (FD_prop$fdis - FD_prop$fdis_mean) / FD_prop$fdis_sd
FD_prop$feve_SES <- (FD_prop$feve - FD_prop$feve_mean) / FD_prop$feve_sd

# Plot FD SES ------------------------------------------------------------------
plot_prop_frich_SES <- ggplot(FD_prop, aes(1/width, frich_SES, color = as.factor(prop))) +
    geom_line(size = 1) +
    scale_color_manual(values = viridis::viridis(3)) +
    xlab("Environmental intensity (1/width)") +
    ylab("Patch functional richness SES") +
    theme(legend.position = "none")
